Unrecognized States: A Theory of Self-Determination and Foreign Influence

### 2015-11-10 draft.tex
finish rewrite of lemma 4 proof

Since the patron is willing to invest up to $\frac{\lambda + \beta + \mu}{1 -\de} -\mu$, when $\frac{\nu}{1 -\de} \geq \frac{\lambda + \beta + \mu}{1 -\de} -\mu$, player $c$'s willingness to counter the patron's investments is sufficient to prevent the \emph{recognition} outcome. A slightly weaker version of this inequality is stated as Restriction (4) so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 4} \\
The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ} -R_p^{WAR}$. For player $c$ it is $\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de} + R_c^{SQ} -R_c^{WAR}$. As noted above, $R_c^{SQ} =0$. So $c$ is willing to invest up to $\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}$ toward preventing fighting if this quantity is negative (leave the case where it is non-negative until later).

$p$ is willing to invest up to $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive. Thus we must only examine the case where this quantity is positive.

We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu + \beta}{1-\de} -\mu$ as this is most restrictive. In this only case of interest, we have $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu> 0$.

Multiplying Restriction (3) by $-(1-p_{1s})$, we have $-\beta \cdot (1-p_{1s}) \geq -\alpha \cdot (1-p_{1s})$. Using this with Restriction (4), we have $\nu p_{1s} - \beta (1-p_{1s}) \geq \lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta> 0 \ \forall p_{1s}$ and therefore also
\[
  \textstyle -\left(\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de}\right) \geq \frac{\lambda p_{1s} - \alpha (1-p_{1s})+ \mu + \beta}{1 - \de}> 0 \ \forall p_{1s}.
\]

Hence when player $p$'s gain from instigating war is positive, player $c$'s loss is negative and larger in absolute magnitude. Player $c$ will thus counter any such investment by player $p$, deterring the investment in the first place. \hfill $\blacksquare$
\\
\\
{\bf Proof of Proposition 2} \\
Suppose that sanctions are applied by player $c$ in period $t+1$ after a Status Quo equilibrium such as the one supported by the construction in Proposition 1 has been established in period $t$. As the only alteration is a reduction in $Q_{st}$ by the amount of the sanctions, it is immediate that if player $p$ is willing and able to supplement player $s$'s status quo payoffs to counteract sanctions, then player $s$'s incentives to play Status Quo are not disturbed. As in Proposition 1, it is not necessary that the patron make these payments immediately if player $s$'s utility from playing Status Quo is high enough relative to its other options, but the degradation of the status quo payoffs will eventually make these payments necessary in order to sustain a Status Quo equilibrium.

>>>>

Since the patron is willing to invest up to $\frac{\lambda + \beta + \mu}{1 -\de} -\mu$, when $\frac{\nu}{1 -\de} \geq \frac{\lambda + \beta + \mu}{1 -\de} -\mu$, player $c$'s willingness to counter the patron's investments is sufficient to prevent the \emph{recognition} outcome. A slightly weaker version of this inequality is stated as Restriction (4) so that this deterrence dynamic is guaranteed to operate through the war lottery as well, as will be seen in Lemma 4. \hfill $\blacksquare$
\\
\\
\noindent {\bf Proof of Lemma 4} \\
The difference in continuation values from switching from the status quo to taking the war lottery for player $p$ is $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ} -R_p^{WAR}$. For player $c$ it is $\frac{-\nu p_{1s} + \beta (1-p_{1s})}{1 -\de} + R_c^{SQ} -R_c^{WAR}$. As noted above, $R_c^{SQ} =0$. So $c$ is willing to invest up to $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}$ toward preventing fighting.

The patron is willing to invest up to $\frac{\lambda p_{1s} - \alpha (1-p_{1s})}{1 -\de} +R_p^{SQ}$ toward instigating fighting when this quantity is positive, and will not invest when the quantity is non-positive. Thus we must only examine the case where this quantity is positive.

We will again set $R_p^{SQ}$ at its largest value of $\frac{\mu + \beta}{1-\de} -\mu$ as this is most restrictive. In this only case of interest, we have $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu> 0$ for an upper bound on the amount the patron is willing to invest to instigate fighting.

If $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  < \omega_{st}\left(R_{pt}\right) - Q_{st}$, the patron's investment in $\omega_{st}$ creates a large enough difference between the payoffs from fighting and the status quo so that player $c$ will not find it worthwhile to re-establish the status quo outcome.

On the other hand, player $c$ would be willing to counter whatever investment player $p$ makes to try to induce fighting---e.g. counter an addition to $\omega_{st}$ with an investment to $q_{st}$---as long as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq \omega_{st}\left(R_{pt}\right) - Q_{st}$.

This can be expanded as $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  \geq -\zeta_{s}(1-\de) + (W_{s,t-1}+R_{pt})p_{1s} + L_{s,t-1}(1-p_{1s}) - Q_{s,t-1} +\mu = \omega_{s,t-1} - Q_{s,t-1} +R_{pt} p_{1s} + \mu$.

Restriction (1) combined with equilibrium play implies that  $Q_{s,t-1} \geq \omega_{s,t-1}$ so a tighter condition for preventing player $p$ from inducing fighting is $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de} - \mu \geq R_{pt} p_{1s}$.

Since $R_{pt} \leq \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} -\mu$, if we can show $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1 -\de}  - \mu \geq p_{1s} \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1 - \de} - p_{1s} \mu$, we will have the guarantee that player $c$ will counterinvest after any effort by the patron to instigate fighting, thereby making $p$'s optimal investment in fighting zero.

Multiplying Restriction (3) by $-(1-p_{1s})$, we have $\frac{-\beta \cdot (1-p_{1s})}{1-\de} -\mu (1-p_{1s}) \geq \frac{-\alpha \cdot (1-p_{1s})}{1-\de}$. Using this with Restriction (4) multiplied by $\frac{1}{1-\de}$, we have $\frac{\nu p_{1s} - \beta (1-p_{1s})}{1-\de} -\mu \geq \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1-\de} -p_{1s} \mu$. Since it is assumed that the war lottery probabilities are non-degenerate, we also have $\frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1-\de} -p_{1s} \mu > p_{1s} \frac{\lambda p_{1s} - \alpha (1-p_{1s}) + \mu + \beta}{1-\de} -p_{1s} \mu$. \hfill $\blacksquare$
\\
\\
{\bf Proof of Proposition 2} \\
Suppose that sanctions are applied by player $c$ in period $t+1$ after a Status Quo equilibrium such as the one supported by the construction in Proposition 1 has been established in period $t$. As the only alteration is a reduction in $Q_{st}$ by the amount of the sanctions, it is immediate that if player $p$ is willing and able to supplement player $s$'s status quo payoffs to counteract sanctions, then player $s$'s incentives to play Status Quo are not disturbed. As in Proposition 1, it is not necessary that the patron make these payments immediately if player $s$'s utility from playing Status Quo is high enough relative to its other options, but the degradation of the status quo payoffs will eventually make these payments necessary in order to sustain a Status Quo equilibrium.
